Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment

### 2020-08-10 manuscript/manuscript.Rmd
osf link adjustments
fixed some typos

Difficulty establishing the analytic reproducibility of research reports has been encountered in several scientific domains, including economics, political science, behavioural ecology, and psychology [@chang2015; @eubank2016; @stodden2018; @obels2019; @hardwicke2018a; @minocher2020; but see @naudet2018]. A preliminary obstacle for many such studies is that research data are typically unavailable [@hardwicke2018; @hardwicke2020a; @wicherts2006]. Even when data can be accessed, suboptimal data management and inadequate documentation can complicate data reuse [@hardwicke2018a; @kidwell2016]. These failures to adequately preserve and share earlier stages of the research pipeline typically prevent downstream assessment of analytic reproducibility.

@@

We employed a retrospective non-comparative case-study design based on @hardwicke2018a. The numerical discrepancy between **original values** reported in the target articles and **reanalysis values** obtained in our reproducibility checks was quantified using percentage error[^percentError] ($PE = \frac{|reanalysis-original|}{original}\times100$). Numerical discrepancies were classified as ‘minor’ (0% > PE < 10%) or ‘major’ (PE >= 10%). If an old *p* value fell on the opposite side of the $\alpha = .05$ boundary relative to a new *p* value we additionally recorded a ‘decision error’. We recorded if there was insufficient information to complete aspects of the analyses.

@@

After author involvement, `r majorErrors` major numerical discrepancies remained amongst the `r valuesChecked` target values examined across all articles (`r paste0(majorErrors_percent,'% ',majorErrors_percentCI)`). This included `r decisionErrors` decision error for which we obtained a statically significant p value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(obtainedValue)` in contrast to the reported non-significant p-value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(reportedValue)`. A scatterplot illustrating the consistency of original and reanalysis p-values is displayed in Figure \@ref(fig:scatterPlot). The frequency of numerical discrepancies by value types is displayed in Figure \@ref(fig:valueTypePlot).

```{r valueTypePlot, fig.cap="Frequency of reproducibility outcomes by value type. Variation/uncertainty measures include standard deviations, standard errors, and confidence intervals. Effect sizes include Cohen's d, Person's r, partial eta squared, and phi. Test statistics include t, F, and chi squared. Central tendency measures include means and medians.", fig.path="figs/"}
source(here('analysis', 'plot_valueType.R')) # build plot
valueTypePlot # show plot
```

@@

Whilst the current findings are concerning, non-reproducibility is fortunately a solvable problem, at least in principle. Journals are well-situated to influence practices related to research transparency; for example, journal data sharing mandates have been associated with marked increases in data availability [@hardwicke2018a; @nuijten2017]. Further requirements to share analysis scripts may also enhance reproducibility as veridical documentation of the analysis process is often poorly captured in verbal prose [@hardwicke2018a]. However, as with data sharing, mere availability may not be sufficient to confer the potential benefits of analysis scripts. The utility of scripts will likely depend on a range of factors, including clarity, structure, and documentation, as well the programming language that is used and whether it is stored in a proprietary format or requires special expertise to understand. In the present study, script availability appeared to offer modest benefits in terms of reproducibility outcomes and time expenditure; however, as only six articles shared scripts[^psychsci_scripts], it is difficult to draw strong conclusions about their impact from this evidence alone. Journals could also conduct independent assessment of analytic reproducibility prior to publication, as has been adopted by the American Journal of Political Science [@jacoby2017]; however, this would naturally require additional resources [for discussion see @sakaluk2014]. Ideally, initiatives intended to improve analytic reproducibility should undergo empirical scrutiny in order to evaluate costs and benefits and identify any policy shortfalls or unintended consequences [@hardwicke2020b].

@@

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on the Open Science Framework on October 18, 2017 (https://osf.io/2cnkq/). All deviations from this protocol or additional exploratory analyses are explicitly acknowledged in Supplementary Information \@ref(prereg). All data exclusions and measures conducted during this study are reported. All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/tfd8z/). To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr [@R-knitr] and papaja [@R-papaja], and is available in a Code Ocean container (https://doi.org/10.24433/CO.1796004.v1) which re-creates the software environment in which the original analyses were performed. Analysis code, reports, and Code Ocean containers are also available for each reproducibility check (Supplementary Information \@ref(vignettes)).

>>>>

Difficulty establishing the analytic reproducibility of research reports has been encountered in several scientific domains, including economics, political science, behavioural ecology, and psychology [@chang2015; @eubank2016; @stodden2018; @obels2019; @hardwicke2018a; @minocher2020; but see @naudet2018]. A preliminary obstacle for many such studies is that research data are typically unavailable [@hardwicke2018; @hardwicke2020a; @wicherts2006]. Even when data can be accessed, suboptimal data management and inadequate documentation can complicate data re-use [@hardwicke2018a; @kidwell2016]. These failures to adequately preserve and share earlier stages of the research pipeline typically prevent downstream assessment of analytic reproducibility.

@@

We employed a retrospective non-comparative case-study design based on @hardwicke2018a. The numerical discrepancy between **original values** reported in the target articles and **reanalysis values** obtained in our reproducibility checks was quantified using percentage error[^percentError] ($PE = \frac{|reanalysis-original|}{original}\times100$). Numerical discrepancies were classified as ‘minor’ (0% > PE < 10%) or ‘major’ (PE >= 10%). If an old p-value fell on the opposite side of the $\alpha = .05$ boundary relative to a new p-value we additionally recorded a ‘decision error’. We recorded when there was insufficient information to complete aspects of the analyses.

@@

After author involvement, `r majorErrors` major numerical discrepancies remained amongst the `r valuesChecked` target values examined across all articles (`r paste0(majorErrors_percent,'% ',majorErrors_percentCI)`). This included `r decisionErrors` decision error for which we obtained a statically significant p-value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(obtainedValue)` in contrast to the reported non-significant p-value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(reportedValue)`. A scatterplot illustrating the consistency of original and reanalysis p-values is displayed in Figure \@ref(fig:scatterPlot). The frequency of numerical discrepancies by value types is displayed in Figure \@ref(fig:valueTypePlot).

```{r valueTypePlot, fig.cap="Frequency of reproducibility outcomes by value type. Variation/uncertainty measures include standard deviations, standard errors, and confidence intervals. Effect sizes include Cohen's d, Pearson's r, partial eta squared, and phi. Test statistics include t, F, and chi squared. Central tendency measures include means and medians.", fig.path="figs/"}
source(here('analysis', 'plot_valueType.R')) # build plot
valueTypePlot # show plot
```

@@

Whilst the current findings are concerning, non-reproducibility is fortunately a solvable problem, at least in principle. Journals are well-situated to influence practices related to research transparency; for example, journal data sharing mandates have been associated with marked increases in data availability [@hardwicke2018a; @nuijten2017]. Further requirements to share analysis scripts may also enhance reproducibility as veridical documentation of the analysis process is often poorly captured in verbal prose [@hardwicke2018a]. However, as with data sharing, mere availability may not be sufficient to confer the potential benefits of analysis scripts. The utility of scripts will likely depend on a range of factors, including clarity, structure, and documentation, as well the programming language that is used and whether they are stored in a proprietary format or require special expertise to understand. In the present study, script availability appeared to offer modest benefits in terms of reproducibility outcomes and time expenditure; however, as only six articles shared scripts[^psychsci_scripts], it is difficult to draw strong conclusions about their impact from this evidence alone. Journals could also conduct independent assessment of analytic reproducibility prior to publication, as has been adopted by the American Journal of Political Science [@jacoby2017]; however, this would naturally require additional resources [for discussion see @sakaluk2014]. Ideally, initiatives intended to improve analytic reproducibility should undergo empirical scrutiny in order to evaluate costs and benefits and identify any policy shortfalls or unintended consequences [@hardwicke2020b].

@@

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on the Open Science Framework on October 18, 2017 (https://osf.io/2cnkq/). All deviations from this protocol or additional exploratory analyses are explicitly acknowledged in Supplementary Information \@ref(prereg). We report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study. All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/n3dej/). To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr [@R-knitr] and papaja [@R-papaja], and is available in a Code Ocean container (https://doi.org/10.24433/CO.1796004.v1) which re-creates the software environment in which the original analyses were performed. Analysis code, reports, and Code Ocean containers are also available for each reproducibility check (Supplementary Information \@ref(vignettes)).

### 2020-08-10 manuscript/supplementary_information.Rmd
typo in perma link
fixed some typos

There was one major numerical discrepancy where a p value was reported as = .001 but we obtained 3.695211e-13. It seems very likely the intention was to report p <.001 as the reported t-value is consistent with this significance level (the authors have confirmed over email that it was probably a typo). This single observed discrepancy is unlikely to be consequential for the article's original conclusions.

@@

R Markdown report: https://perma.cc/K5LB-WD45s

@@

Two major numerical discrepancies with p-values remained. In both cases, the authors reported that p = .001, but the values we obtained were lower than this. We discussed this issue with the authors and they said that it is their policy to report p values that are <.001 as p = .001. Unfortunately this approach guarantees inaccurate reporting. According to the APA, researchers should "report p values less than .001 as p < .001" (APA, 2009; p. 114). The reproducibility issues we encountered do not appear to undermine the conclusions drawn in the original article.

@@

However, we continued to have difficulties with the section reporting post-hoc comparisons. There were three major numerical discrepancies for p-values. The authors report using 'Sidak corrections' but it is unclear if they were applied to the alpha threshold (correctly) or the p-values themselves. If corrections were applied to the p values, then this could potentially explain the differences, but the article does not identify the family of hypotheses for which corrections were applied. So we also cannot correct the alpha threshold. We have decided to report the major numerical discrepancies for the p values, but also record an insufficient information error as we cannot determine whether these were decision errors or not. Additionally, for one degree of freedom, we obtained 4 whereas the article reports 3. This may be a typo but the cause is not clear.

>>>>

There was one major numerical discrepancy where a p-value was reported as = .001 but we obtained 3.695211e-13. It seems very likely the intention was to report p <.001 as the reported t-value is consistent with this significance level (the authors have confirmed over email that it was probably a typo). This single observed discrepancy is unlikely to be consequential for the article's original conclusions.

@@

R Markdown report: https://perma.cc/K5LB-WD45

@@

Two major numerical discrepancies with p-values remained. In both cases, the authors reported that p = .001, but the values we obtained were lower than this. We discussed this issue with the authors and they said that it is their policy to report p-values that are <.001 as p = .001. Unfortunately this approach guarantees inaccurate reporting. According to the APA, researchers should "report p-values less than .001 as p < .001" (APA, 2009; p. 114). The reproducibility issues we encountered do not appear to undermine the conclusions drawn in the original article.

@@

However, we continued to have difficulties with the section reporting post-hoc comparisons. There were three major numerical discrepancies for p-values. The authors report using 'Sidak corrections' but it is unclear if they were applied to the alpha threshold (correctly) or the p-values themselves. If corrections were applied to the p-values, then this could potentially explain the differences, but the article does not identify the family of hypotheses for which corrections were applied. So we also cannot correct the alpha threshold. We have decided to report the major numerical discrepancies for the p-values, but also record an insufficient information error as we cannot determine whether these were decision errors or not. Additionally, for one degree of freedom, we obtained 4 whereas the article reports 3. This may be a typo but the cause is not clear.
